\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark

### 2014-08-22 ResponseToReview.tex

\documentclass[11pt]{article}
\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
\newcommand{\res}[1]{\noindent \textcolor{blue}{{#1}} \\}
\newcommand{\jri}[1]{\noindent \textcolor{red}{{#1}} \\}
\newcommand{\lev}[1]{\noindent \textcolor{green}{{#1}} \\}
\newcommand{\mbh}[1]{\noindent \textcolor{Dandelion}{{#1}}\\}
\usepackage{lipsum}
\usepackage{url}
\usepackage[usenames,dvipsnames]{color}
\usepackage{multirow}
\renewcommand{\baselinestretch}{1.5}
\usepackage{natbib} % package for formatting citations
\bibliographystyle{AJB} %indicating bibliography uses the AJB.bst file for

\begin{document}

\noindent Dear Dr. Schnabel and editorial staff at the \emph{American Journal of Botany}, \\

@@

\noindent Jeffrey Ross-Ibarra and Matthew Hufford

\section*{Response to Associate Editor:}

I thought that the neighbor-joining analysis was of little value, because it was not clear to me just how the result could be expected to provide information about Hopscotch genotype.  Why would you expect individuals with similar Hopscotch genotypes to cluster together?  Also, although you don't provide any information on the distribution of variable sites among accessions and taxa, your data set of only 48 and 40 segregating sites is almost certainly not sufficiently variable to resolve relationships among the very large number of accessions you were analyzing.  As a result, you produce mostly a huge polytomy, with a small amount of structure that appears to reflect the genetic similarities and differences already evident in Table 1.  The trees themselves are difficult for a reader to interpret, because none of the accession names at the branch tips (other than TIL) are defined in the text.\\

\res{We agree entirely with this assessment of the Neighbor Joining trees. 
Our intention with the NJ trees was to show that there is no clear signal of selection on the \emph{Hopscotch} genotype across sequence in the \emph{tb1} genomic region.
Presumably, if the \emph{Hopscotch} element were under strong selection in teosinte populations we would expect to see \emph{tb1} haplotype groups in an NJ tree due to linkage disequilibrium and Hill-Robertson effects.
In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch}. \mbh{CITE} 
We agree with the reviewer that the trees could be confusing and difficult to interpret, and the large polytomies do not provide information beyond our assertion that there is not strong haplotype structure broadly in the \emph{tb1} genomic region that correlates with the \emph{Hopscotch} genotype.
We have therefore opted to keep the NJ trees in the supplement where interested readers can view them but where they don't unnecessarily complicate our main story.
In addition, we have added explanation of the accession names in the supplemental figure legend.}

@@

\jri{need to add explanation of accession names in separate table or in figure legend.} \mbh{I'd add this to the legend}

Phenotyping experiments - These are curious experiments.  First, why did you not generate a population of individuals with known genotypes and then test for the effect of genotype on phenotype?  Why take a random sample and just hope that you will get the necessary number of individuals of each genotype?  How can you randomize the growing conditions with respect to genotype, if you don't know genotypes ahead of time but instead determine them after all the growing measurements are taken?\\

\res{We agree that a planned grow-out with previously determined genotypes could be useful.  
However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, \mbh{CITE} and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
Moreover, our sampling of individuals from this population for our grow out was not random.  
We initially genotyped a single individual from each sampling site within the population and then targeted our sample for the grow out to include individuals from sampling sites where individuals were homozygous or heterozygous for the \emph{Hopscotch} allele.
Our reasoning in using this approach was that we wanted a high proportion of both \emph{Hopscotch} positive and \emph{Hopscotch} negative individuals for phenotypic comparisons and, given that the \emph{Hopscotch} allele is typically rare in teosinte, we wanted to enrich our sample for this allele. 
We were overly successful in this approach and had a higher frequency of individuals with the \emph{Hopscotch} positive allele.
We have clarified our sampling methodology on page \mbh{X}, lines \mbh{XX}.
Finally, plants were situated randomly in the greenhouse prior to genotyping.
Regardless of the fact that genotypes were eventually determined, this still represents a randomized experimental design.}

@@

Why are sample sizes not reported?  You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class.\\

\res{We have reported sample size information for all three genotypes on page \mbh{X} lines \mbh{XX}. We have also corrected our typo for sample size and the Results and Methods sections are now consistent.} \jri{I didn't see sample sizes for each of the genotpyes. are they there? if not, please add.}

@@

It was not clear to me what ``to detect the observed effect" means in lines 12-13 of p. 9.  Your Phenotyping 1 results show no effect of genotype on tillering.  Thus, it is not clear what effect you are referring to.\\

What was the purpose of the Phenotyping 2  experiment?  You already showed no effect of genotype on phenotype using populations where you are most likely to see segregation for the insertion, so it was not clear to me why sampling more broadly, especially from populations with no evidence of the Hopscotch insertion, would be expected to improve your understanding of this relationship.\\

\res{On Day 40 of the Phenotyping 1 experiment we saw a weak positive correlation between the \emph{Hopscotch} insertion and tillering index (p=0.0848).  
Using data from Phenotyping 1, we conducted a power analysis to determine the number of individuals we would need in both homozygous genotypic classes to observe a significant result.
This analysis indicated we would need 71 individuals in both classes to detect a significant difference in tillering index.  
We attempted to sample enough individuals from the San Lorenzo population (the same population sampled for Phenotyping 1) where we had already shown \emph{Hopscotch} to be segregating at high frequency in order to detect a significant difference between homozygous genotypes.  
Our Phenotyping 2 experiment detects the same trend reported in Phenotyping 1, but the trend is still not significant.
We agree with the editor that this second, non-significant result adds little to the manuscript and we have thus removed Phenotyping 2 and refer to a single grow-out in our revised version of the manuscript.}

@@

\mbh{I don't think the next paragraph is necessary}

\res{Note that all the sampling for both phenotyping experiments was limited to a single large population from which we had a large number of seed collected. Seed was collected from various "sampling sites" within the contiguous population at which we had estimated Hopscotch allele frequency. For our more broad genotyping of both landrace maize, ssp. \emph{parviglumis}, and ssp. \emph{mexicana} we sampled as many populations as we could. For most of these populations the prevalence of the Hopscotch was previously unknown, and we were interested in having an idea of the overall distribution and frequency of the Hopscotch throughout these locations.}

The discussion of introgression, genetic drift, and selection in your Discussion section seems to lack coherence.  It sometimes focuses on explaining the unexpectedly high frequency of Hopscotch in present-day populations, whereas other times it seems to be addressing the presence or absence of Hopscotch in teosinte in general, and it also sometimes seems to be addressing selection on the tb1 locus.

\res{We have edited this section for clarity to better emphasize our main point that the \emph{Hopscotch} allele, which is known to produce a domesticated phenotype of reduced tiller number, is at higher than expected frequency in a subset of teosinte populations and may play an adaptive role in certain environments encountered by teosinte.  
Surprisingly, based on our grow-out of individuals from the San Lorenzo population, tiller number is not a major phenotypic difference between teosinte plants with and without the \emph{Hopscotch} insertion, perhaps indicating the insertion affects multiple phenotypes in teosinte}

@@

\section*{Response to Reviewer 1}

Reviewer \#1: In their paper `Natural variation in teosinte at the domestication locus teosinte branched1 (tb1)', Vann et al survey a large sample of maize and teosinte individuals for the presence of the Hopscotch transposable element upstream of tb1. The authors find that Hopscotch is more widespread, across a large sampled area in Mexico, than previously thought, particularly in parviglumis. The authors go on to sequence regions up and downstream of Hopscotch in subset of their initial sample. In analyzing their sequence data, they find no evidence for recent introgression of the maize Hopscotch locus into wild teosintes. They also find evidence for selection acting on tb1 in parviglumis. In greenhouse experiments, no difference in tiller index or tiller number was observed between teosinte lines with or without Hopscotch.

They present an interesting result, their conclusions are supported by their data, and the methods they use are appropriate. However, I feel like they need to make some further effort to understand what is happening mechanistically. What is Hopscotch doing in maize vs these sampled teosintes? I feel like the expression analyses and phenotyping experiments they suggest should be part of a later study should be part of this study. At a minimum, qRT-PCR experiments should be performed assessing tb1 expression in their with/without Hopscotch lines. Expression analyses of gt1, te1, and tru1 would be great, since they invoke variation at these loci to explain their results.

\res{We attempted to measure expression of \emph{tb1} following the protocol of \citep{hubbard2002expression} but were unable to get reasonable results. 
Subsequent discussion with colleagues revealed teosinte expression of \emph{tb1} is very difficult to detect except in early ear tissue.}

\mbh{This reads as a pretty weak response...we tried to measure expression, we got weird results that we don't report, and our colleagues tells us this is hard.  I think I'd lean toward leaving this out and just reemphasizing that expression experiments are beyond the scope of the current work}

\res{Extensive expression analysis of individuals with and without the \emph{Hopscotch} insertion at \emph{tb1} and a number of other genes is a good suggestion for follow-up work but would be a substantial undertaking and indeed more appropriate for another paper. }

@@

Another potential avenue, that the authors also suggest, lies in more directed and exhaustive greenhouse experiments. Variation in red and far-red light has been shown to be important in regulating both tb1 (in Sorghum) and gt1 in maize. Growth chamber experiments (with or without supplemental far red light, with or without Hopscotch), coupled to expression analyses in seedlings, seem like they should be possible.  In the greenhouse experiments that were done, phenotyping was incomplete. Apart from tillering, Hopscotch and natural variation at the tb1 locus appears to affect ear architecture and internode length. Was there any evidence for variation in these other traits? In the longer term, it would be nice to see what happens when you introgress these particular parviglumis loci into maize.\\

\res{We did not measure ear architecture or internode length. 
While \emph{tb1} does impact these traits, our primary interest was in the effect of \emph{tb1} on tillering as a potentially ecologically relevant phenotype given its known substantial effects on this trait in a maize background.
Both additional greenhouse and expression analyses would indeed be useful avenues of further investigation.
Nonetheless, these are extensive additional experiments which we are unable to include at this time. 
We have, however, added text to the conclusion (\mbh{Page X, Lines X}) of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte:
``Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g. \emph{gt1}, \emph{tru1}, and \emph{te1}) as well as include a more exhaustive phenotyping including all traits.  
These data, in conjunction with more exhaustive phenotyping, should help reveal the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte."}

@@

One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown.

\res{We have added the Balsas River in a revised version of Figure 1.}

### 2014-08-22 ResponseToReview.tex

\noindent Jeffrey Ross-Ibarra and Matthew Hufford

\section*{Response to Associate Editor:}

I thought that the neighbor-joining analysis was of little value, because it was not clear to me just how the result could be expected to provide information about Hopscotch genotype.  Why would you expect individuals with similar Hopscotch genotypes to cluster together?  Also, although you don't provide any information on the distribution of variable sites among accessions and taxa, your data set of only 48 and 40 segregating sites is almost certainly not sufficiently variable to resolve relationships among the very large number of accessions you were analyzing.  As a result, you produce mostly a huge polytomy, with a small amount of structure that appears to reflect the genetic similarities and differences already evident in Table 1.  The trees themselves are difficult for a reader to interpret, because none of the accession names at the branch tips (other than TIL) are defined in the text.\\

\res{We agree entirely with this assessment of the Neighbor Joining trees. 
Our intention with the NJ trees was to show that there is no clear signal of selection on the \emph{Hopscotch} genotype across sequence in the \emph{tb1} genomic region.
Presumably, if the \emph{Hopscotch} element were under strong selection in teosinte populations we would expect to see \emph{tb1} haplotype groups in an NJ tree due to linkage disequilibrium and Hill-Robertson effects.
In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch} (Studer \emph{et al.} 2011).
We agree with the reviewer that the trees could be confusing and difficult to interpret, and the large polytomies do not provide information beyond our assertion that there is not strong haplotype structure broadly in the \emph{tb1} genomic region that correlates with the \emph{Hopscotch} genotype.
We have therefore opted to keep the NJ trees in the supplement where interested readers can view them but where they don't unnecessarily complicate our main story.
In addition, we have added explanation of the accession names in the supplemental figure legend.}

### 2014-08-22 ResponseToReview.tex

One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown.

\res{We have added the Balsas River in a revised version of Figure 1.}

\section*{Response to Reviewer 2} 
Although this paper brings up an interesting question regarding the role of tb1 standing variation in teosinte, it is not able to answer it completely.  It presents results that have been published previously about the presence of hopscotch in teosinte, maize, and mexicana.

\res{Reviewer 2 is correct in stating that it was known that the Hopscotch was present in some teosinte individuals \mbh{CITE}; however, previous studies sampled a small fraction of the populations and individuals per population that we present here and thus did not provide comprehensive estimates. 
Our study is not only concerned with the presence or absence of the \emph{Hopscotch} allele in teosinte but also the frequency at which it segregates in a range-wide sample of teosinte.  
This finer level of detail sheds light on the evolutionary history of the \emph{Hopscotch} insertion (\emph{i.e.}, the domestication allele at \emph{tb1}) in teosinte. 
We have added clarifying text to the introduction: ``The effects of the \emph{Hopscotch} insertion have been studied in maize (Studer \emph{et al.} 2011), and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes (Studer and Doebley 2012), but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte. Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of ssp. \emph{parviglumis}, ssp. \emph{mexicana}, and landrace maize; yet little is known about the frequency at which the \emph{Hopscotch} is segregating in natural populations."}

@@

Nevertheless, the manuscript is worth publishing as it does present data that may help rule out some possible explanations as to why tb1 is so common in parviglumus.\\

\res{Thank you; we hope that our paper in conjunction with future research and experiments can help shed light on the ecological significance of both \emph{tb1} and the \emph{Hopscotch} insertion in teosinte.}

@@

This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied.\\

\jri{tweak intro and abstract, add a bit to discussion maybe, explain text changes here.}
\lev{Jeff and Matt, this is already in the intro. I don't see what more we can say that would be appropriate. ``The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte."}
\jri{add some discussion about domesitcaiton traits in other species, weedy rice, etc. }

\mbh{Yes, I think we could include more text on traits commonly under selection during domestication (\emph{i.e.}, the domestication syndrome) and perhaps highlight how these are relevant to the fitness of wild relatives and could have been the targets of natural selection prior to domestication.  In my opinion there is a false dichotomy out there regarding "natural" vs. "artificial" selection.  Give this a shot, LV, and I'll edit with my POV.}

@@

Methods and statistics were appropriately used as far as my expertise allows me to judge.
Specific suggestions follow:

Introduction:
Page 4, Lines 7 - 9: You say "Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic eﬀects has become increasingly useful to crop improvement (Kovach and McCouch, 2008).".  Can you tell us why an understanding is useful? Can you give some examples?\\

\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and flowering time into domesticated rice (Kovach and McCouch 2008)."} \jri{read matt's teosinte TIG paper, there are good cites there about use of teosinte specifically.}

\mbh{lots of fodder in this manuscript as well, LV: \url{http://bioscience.oxfordjournals.org/content/61/7/559.full}}

@@

8) Page 5, line 12 - 14: You say "The tb1 locus appears to play an important role in the shade avoidance pathway in Zea mays and other grasses and may therefore be crucial to the ecology of teosinte (Kebrom and Brutnell, 2007; Lukens and Doebley, 1999)." Please clarify; do the references belong to the first half of the sentence, supporting tb1 playing a role in Zea mays, or the second half of the sentence, supporting tb1 possibly being crucial to the ecology of teosinte? Which of the two distinct things have actually been studied? If the first, could you tell us how tb1 has been shown to work in maize under shade conditions? If the second, was it actually studied, or merely hypothesized in teosinte?\\

\res{We have moved the references cited here to more clearly reflect the papers' findings.  We have also added text describing the role of \emph{tb1} in the shade avoidance pathway and clarifying the manner in which this could affect fitness in teosinte}

\mbh{References need to follow text saying \emph{tb1} is important in the shade avoidance pathway and we need to add a sentence or two clarifying what it does in this pathway and the implications for teosinte fitness}

9) Page 5, Lines 15 - 16: you say you ``aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic effects of the insertion in parviglumis."  The mexicana seem a bit tacked on and not much discussed. Did you phenotype mexicana? Can you present the data here, if so? Does the analysis of this group of teosintes add anything to your conclusions?\\

### 2014-08-22 ResponseToReview.tex

\res{We agree entirely with this assessment of the Neighbor Joining trees. 
Our intention with the NJ trees was to show that there is no clear signal of selection on the \emph{Hopscotch} genotype across sequence in the \emph{tb1} genomic region.
Presumably, if the \emph{Hopscotch} element were under strong selection in teosinte populations we would expect to see \emph{tb1} haplotype groups in an NJ tree due to linkage disequilibrium and Hill-Robertson effects.
In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch} (Studer \emph{et al.} 2011).
We agree with the reviewer that the trees could be confusing and difficult to interpret, and the large polytomies do not provide information beyond our assertion that there is not strong haplotype structure broadly in the \emph{tb1} genomic region that correlates with the \emph{Hopscotch} genotype.
We have therefore opted to keep the NJ trees in the supplement where interested readers can view them but where they don't unnecessarily complicate our main story.
In addition, we have added explanation of the accession names in the supplemental figure legend.}

Phenotyping experiments - These are curious experiments.  First, why did you not generate a population of individuals with known genotypes and then test for the effect of genotype on phenotype?  Why take a random sample and just hope that you will get the necessary number of individuals of each genotype?  How can you randomize the growing conditions with respect to genotype, if you don't know genotypes ahead of time but instead determine them after all the growing measurements are taken?\\

\res{We agree that a planned grow-out with previously determined genotypes could be useful.  
However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, (Studer \emph{et al.} 2011) and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
Moreover, our sampling of individuals from this population for our grow out was not random.  
We initially genotyped a single individual from each sampling site within the population and then targeted our sample for the grow out to include individuals from sampling sites where individuals were homozygous or heterozygous for the \emph{Hopscotch} allele.
Our reasoning in using this approach was that we wanted a high proportion of both \emph{Hopscotch} positive and \emph{Hopscotch} negative individuals for phenotypic comparisons and, given that the \emph{Hopscotch} allele is typically rare in teosinte, we wanted to enrich our sample for this allele. 
We were overly successful in this approach and had a higher frequency of individuals with the \emph{Hopscotch} positive allele.
We have clarified our sampling methodology on page \mbh{X}, lines \mbh{XX}. Finally, plants were situated randomly in the greenhouse prior to genotyping.
Regardless of the fact that genotypes were eventually determined, this still represents a randomized experimental design.}

### 2014-08-22 ResponseToReview.tex

\res{We agree that a planned grow-out with previously determined genotypes could be useful.  
However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, (Studer \emph{et al.} 2011) and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
Moreover, our sampling of individuals from this population for our grow out was not random.  
We initially genotyped a single individual from each sampling site within the population and then targeted our sample for the grow out to include individuals from sampling sites where individuals were homozygous or heterozygous for the \emph{Hopscotch} allele.
Our reasoning in using this approach was that we wanted a high proportion of both \emph{Hopscotch} positive and \emph{Hopscotch} negative individuals for phenotypic comparisons and, given that the \emph{Hopscotch} allele is typically rare in teosinte, we wanted to enrich our sample for this allele. 
We were overly successful in this approach and had a higher frequency of individuals with the \emph{Hopscotch} positive allele.
We have clarified our sampling methodology on page \mbh{X}, lines \mbh{XX}. Finally, plants were situated randomly in the greenhouse prior to genotyping.
Regardless of the fact that genotypes were eventually determined, this still represents a randomized experimental design.}

Why are sample sizes not reported?  You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class.\\

\res{We have reported sample size information for all three genotypes on page \mbh{X} lines \mbh{XX}. \lev{will fix the line numbers when we've made sure the manuscript text is okay to go}. We have also corrected our typo for sample size and the Results and Methods sections are now consistent.} \jri{I didn't see sample sizes for each of the genotpyes. are they there? if not, please add.} \lev{they should be in the results section.}

@@

It was not clear to me what ``to detect the observed effect" means in lines 12-13 of p. 9.  Your Phenotyping 1 results show no effect of genotype on tillering.  Thus, it is not clear what effect you are referring to.\\

What was the purpose of the Phenotyping 2  experiment?  You already showed no effect of genotype on phenotype using populations where you are most likely to see segregation for the insertion, so it was not clear to me why sampling more broadly, especially from populations with no evidence of the Hopscotch insertion, would be expected to improve your understanding of this relationship.\\

\res{On Day 40 of the Phenotyping 1 experiment we saw a weak positive correlation between the \emph{Hopscotch} insertion and tillering index (p=0.0848).  
Using data from Phenotyping 1, we conducted a power analysis to determine the number of individuals we would need in both homozygous genotypic classes to observe a significant result.
This analysis indicated we would need 71 individuals in both classes to detect a significant difference in tillering index.  
We attempted to sample enough individuals from the San Lorenzo population (the same population sampled for Phenotyping 1) where we had already shown \emph{Hopscotch} to be segregating at high frequency in order to detect a significant difference between homozygous genotypes.  
Our Phenotyping 2 experiment detects the same trend reported in Phenotyping 1, but the trend is still not significant.
We agree with the editor that this second, non-significant result adds little to the manuscript and we have thus removed Phenotyping 2 and refer to a single grow-out in our revised version of the manuscript.}

\mbh{I don't think the next paragraph is necessary} \lev{which next one? the one that starts 'Note that all the sampling for both...}

### 2014-08-25 ResponseToReview.tex

\res{We agree entirely with this assessment of the Neighbor Joining trees. 
Our intention with the NJ trees was to show that there is no clear signal of selection on the \emph{Hopscotch} genotype across sequence in the \emph{tb1} genomic region.
Presumably, if the \emph{Hopscotch} element were under strong selection in teosinte populations we would expect to see \emph{tb1} haplotype groups in an NJ tree due to linkage disequilibrium and Hill-Robertson effects.
In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch} (Studer \emph{et al.} 2011).
We agree with the reviewer that the trees could be confusing and difficult to interpret, and the large polytomies do not provide information beyond our assertion that there is not strong haplotype structure broadly in the \emph{tb1} genomic region that correlates with the \emph{Hopscotch} genotype.
We have therefore opted to keep the NJ trees in the supplement where interested readers can view them but where they don't unnecessarily complicate our main story.
In addition, we have added explanation of the accession names in the supplemental figure legend.}

Phenotyping experiments - These are curious experiments.  First, why did you not generate a population of individuals with known genotypes and then test for the effect of genotype on phenotype?  Why take a random sample and just hope that you will get the necessary number of individuals of each genotype?  How can you randomize the growing conditions with respect to genotype, if you don't know genotypes ahead of time but instead determine them after all the growing measurements are taken?\\

\res{We agree that a planned grow-out with previously determined genotypes could be useful.  
However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, (Studer \emph{et al.} 2011) and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
Moreover, our sampling of individuals from this population for our grow out was not random.  
We initially genotyped a single individual from each sampling site within the population and then targeted our sample for the grow out to include individuals from sampling sites where individuals were homozygous or heterozygous for the \emph{Hopscotch} allele.
Our reasoning in using this approach was that we wanted a high proportion of both \emph{Hopscotch} positive and \emph{Hopscotch} negative individuals for phenotypic comparisons and, given that the \emph{Hopscotch} allele is typically rare in teosinte, we wanted to enrich our sample for this allele. 
We were overly successful in this approach and had a higher frequency of individuals with the \emph{Hopscotch} positive allele.
We have clarified our sampling methodology on pages 8-9, lines 174-185. Finally, plants were situated randomly in the greenhouse prior to genotyping.
Regardless of the fact that genotypes were eventually determined, this still represents a randomized experimental design.}

@@

Why are sample sizes not reported?  You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class.\\

\res{We have reported sample size information for all three genotypes on page 11 lines 266-267.}

### 2014-08-25 Supplement_mbh.tex
adding comments to manuscript file

\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S1 -- Page \thepage}}

@@

\pagenumbering{arabic}
\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S2 -- Page \thepage}}
% Appendix S2. Mays sampling

@@

\clearpage
\pagenumbering{arabic}
\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S3 -- Page \thepage}}
% Appendix S3. Cartoon representation of tb1 locus
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=180mm]{FigS1LocusCartoon.pdf}
   \flushleft{Appendix S3: Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red). Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Region 1; within the 5' UTR, and Region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}
  \end{center}
\end{figure*}
%-------------------------------------------------------------------

\clearpage
\pagenumbering{arabic}
\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S4 -- Page \thepage}}

@@

\clearpage
\pagenumbering{arabic}
\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S5 -- Page \thepage}}

% Appendix S5. NJ Trees
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=150mm]{FigS3NJtrees.pdf}
    \flushleft{Appendix S5: Neighbor-joining tree of the sequenced region in the 5' UTR (right; Region 1) and the 66,169 bp upstream region (left; Region 2) of \emph{tb1}. Individuals with genotype data are colored: Homozygous for the teosinte (no \emph{Hopscotch}) allele (red), homozygous for the maize (\emph{Hopscotch}) allele (blue), heterozygotes (purple). TILs (teosinte inbred lines) are colored in green, with stars indicating the 3 TILs known to have the \emph{Hopscotch} insertion. Black indicates individuals not genotyped for the \emph{Hopscotch} insertion.} 
  \end{center}
\end{figure*}
%-------------------------------------------------------------------
\end{document}

>>>>

%\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S1 -- Page \thepage}}

@@

%\pagenumbering{arabic}
%\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S2 -- Page \thepage}}
% Appendix S2. Mays sampling

@@

\clearpage
%\pagenumbering{arabic}
%\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S3 -- Page \thepage}}
% Appendix S3. Cartoon representation of tb1 locus
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=180mm]{FigS1LocusCartoon.pdf}
   \flushleft{Appendix S3: Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red). Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Region 1; within the 5' UTR, and Region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}
  \end{center}
\end{figure*}
%-------------------------------------------------------------------

\clearpage
%\pagenumbering{arabic}
%\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S4 -- Page \thepage}}

@@

\clearpage
%\pagenumbering{arabic}
%\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S5 -- Page \thepage}}

% Appendix S5. NJ Trees
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=150mm]{FigS3NJtrees.pdf}
    \flushleft{Appendix S5: Neighbor-joining tree of the sequenced region in the 5' UTR (right; Region 1) and the 66,169 bp upstream region (left; Region 2) of \emph{tb1}. Individuals with genotype data are colored: Homozygous for the teosinte (no \emph{Hopscotch}) allele (red), homozygous for the maize (\emph{Hopscotch}) allele (blue), heterozygotes (purple). TILs (teosinte inbred lines) are colored in green, with stars indicating the 3 TILs known to have the \emph{Hopscotch} insertion. Black indicates individuals not genotyped for the \emph{Hopscotch} insertion. EjuA refers to individuals from population Ejutla A, EjuB from Ejutla B, SLO from San Lorenzo, and MSA from La Mesa. Remaining individuals are lines of maize (\emph{Zea mays} ssp. \emph{mays}).  
  \end{center}
\end{figure*}
%-------------------------------------------------------------------
\end{document}
